
```{r Code relating to failed analysis on clutch completion and species.}
#Transforming body mass data using log10 and square root to attempt to make data more normally distributed - unsuccessful.
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_log10()
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_sqrt()

ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_log10()
ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_sqrt()

ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_log10()
ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_sqrt()

#Plotting a violin chart to explore whether body mass can explain successful clutch completion.
ggplot(data=penguins_data_clean, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="point", color="black")

#Plotting violin charts for each species to explore relationship between body mass and clutch completion in each species individually.
ggplot(data=adelie_data, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="pointrange", color="black")

ggplot(data=chinstrap_data, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="pointrange", color="black")

ggplot(data=gentoo_data, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="pointrange", color="black")
```

```{r excess code which will probably be deleted -> unsure what is and is not useful and need to sort out}
#Plotting histogram to ensure data follows normal distribution
ggplot(data = penguins_data_clean, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))

```{r possibly excess code relating to clutch completion vs body mass}
#Setting colour-blind friendly colour palette to distinguish data relating to different species, and defining shapes for each sex.
species_colours <- c("Adelie" = "darkorange", "Chinstrap" = "magenta2", "Gentoo" = "blue")
sex_shapes <- c("MALE" = "+", "FEMALE" = "20")

#Making function to create datasets only containing information on species, clutch completion, body mass and sex.
refining_penguins_data <- function(clean_data){
  clean_data %>% 
    select(species, body_mass_g, clutch_completion, sex) %>% 
    remove_NA()}

#Making new, refined dataset which only contains data on penguin species, size, sex and clutch completion.
penguins_data_refined <- refining_penguins_data(penguins_data_clean)
write.csv(penguins_data_refined, here("data", "penguins_data_refined.csv"))
penguins_data_refined <- read.csv(here("data","penguins_data_refined.csv"))
head(penguins_data_refined)

#Plotting violin chart of body mass against clutch completion for all species, without accounting for sex.
ggplot(data = penguins_data_refined, aes(x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color = species, shape = sex), width = 1) + geom_jitter(aes(color = species), alpha = 0.25, width = 0.5) + scale_colour_manual(values = species_colours) + scale_shape_manual(values = sex_shapes)
```

```{r Code related to failed no interaction modelling for ANCOVA of culmen depth, length and species.}

plot(log_bill_model_no_int, which=2)
plot(log_bill_model_no_int, which=1)
summary(log_bill_model_no_int)
anova(log_bill_model_no_int)

plot_no_int <- ggplot(penguins_data_clean, aes(x=culmen_length_mm, y=culmen_depth_mm, color = species)) + geom_point() + geom_line() + scale_colour_manual(values = species_colours) + xlab("Culmen length (mm)") + ylab("Culmen depth (mm)") + theme_bw()

plot_no_int
```

```{r Investigating all species data.}

#Creating histograms of data on culmen length in all species to investigate data and assess distribution of data points, as well as effects of any transformations.
ggplot(data=penguins_data_clean, aes(x=culmen_length_mm)) + geom_histogram()

#Log transforming culmen_length_mm using log10() function, and plotting histogram.
log_length <- log10(penguins_data_clean$culmen_length_mm) 
ggplot(data=penguins_data_clean, aes(x=log_length)) + geom_histogram()

#Square-root transforming culmen_length_mm using sqrt() function, and plotting histogram.
sqrt_length <- sqrt(penguins_data_clean$culmen_length_mm)
ggplot(data = penguins_data_clean, aes(x=sqrt_length)) + geom_histogram()
#Transformations were not that effective, although log transformation did make data slightly less right-skewed; will use log-transformed culmen length data for analysis.

#Creating histograms of data on culmen depth in all species to investigate data and assess distribution of data points, as well as effects of any transformations.
ggplot(data=penguins_data_clean, aes(x=culmen_depth_mm)) + geom_histogram()

#Log transforming culmen_depth_mm using log10() function, and plotting histogram.
log_depth <- log10(penguins_data_clean$culmen_depth_mm)
ggplot(data=penguins_data_clean, aes(x=log_depth)) + geom_histogram()

#Square-root transforming culmen_depth_mm using sqrt() function, and plotting histogram.
sqrt_depth <- sqrt(penguins_data_clean$culmen_depth_mm)
ggplot(data = penguins_data_clean, aes(x=sqrt_depth)) + geom_histogram()
#Transformations were not effective. Will use untransformed culmen depth data for further analysis.
```

```{r Creating exploratory figure}
#Creating an exploratory figure looking at the relationship between culmen depth and culmen length, as well as how this differs between species. 
#  geom_jitter() shows data points on the plot and also ensures they are a specified width so as to not obscure the graph.
#  seed ensures the jitter is generated the same way every time, making the figure more reproducible.
#  scale_colour_manual function allows for different species to be designated the same colours    every time in a colour blind friendly palette, making the figure more accessible and more       reproducible. 
#     Also allows for more in depth exploration of initial data before statistical analysis          conducted.

#Making a colour blind-friendly palette useful for distinguishing species on figures. Specifying this ensures the different species are represented by the consistent colours on any graph, making this document more accessible (colour blind friendly) and more reproducible.
species_colours <- c("Adelie" = "darkorange", 
                     "Chinstrap" = "magenta2", 
                     "Gentoo" = "blue")

#Encodes information for all-species exploratory graph.
bill_size_exploratory <- ggplot(data=penguins_data_clean, 
                         aes(x=log_length, 
                            y=culmen_depth_mm)) + 
                         geom_point(aes(color=species)) + 
                         geom_jitter(aes(color = species), alpha = 0.25, position =                                     position_jitter(width = 0.4, seed=0)) +
                         scale_colour_manual(values = species_colours) + 
                         xlab("Culmen Length (mm)") + 
                         ylab("Culmen Depth (mm)")

#plot() function allows graph to be outputted.
plot(bill_size_exploratory)

#agg_png() saves figure as .png file in figures folder.
agg_png("figures/bill_size_exploratory.png", 
        width = 30,
        height = 15,
        units = "cm",
        res = 300,
        scaling = 1.125)

#print() ensures figure saved in "figures" folder actually has graph encoded within it, as opposed to blank file.
print(bill_size_exploratory)
dev.off()
```

```{r}
#Creating model looking at effect of sex on culmen length in chinstraps without interaction with body mass.
chinstrap_log_bill_model_noint <- lm(culmen_length_mm ~ sex + body_mass_g, chinstrap_data)

#Makes plots to ensure assumptions of analysis not violated.
par(mfrow = c(1,2)) #par() function allows multi-panel figure of the assumptions plots encoded below to be made, allowing assessment of normality and heteroscedascity of model.
plot(chinstrap_log_bill_model_noint, which = 2) # which=2 specifies QQ plot to compare the data in this model to a theoretical normal dataset, to verify data is normally distributed. As points mostly lie on straight line, can assume this sample is normally distributed.
plot(chinstrap_log_bill_model_noint, which = 1) # which=1 specifies graph of residual values vs fitted values, to verify population variances are equal between samples. Both groups seem to have equal distribution above and below line, so can assume assumption is not violated.

#Checking outputs of analysis of model without interaction to assess adjusted R-squared and values for slope and y-intercept of males and females.
summary(chinstrap_log_bill_model_noint)

#Conducting an ANOVA of the model without interaction to assess what the significant effects are.
chinstrap_noint_anova <- aov(chinstrap_log_bill_model_noint)
summary(chinstrap_noint_anova)

#Conducting ANOVA to assess if model without interaction is a better fit than model with interaction (i.e ifthere is a significant interaction between sex and culmen length in chinstrap penguins).
chinstrap_test_fit_diff <- anova(chinstrap_log_bill_model_int, chinstrap_log_bill_model_noint)
chinstrap_test_fit_diff
```

Some stats