
```{r Code relating to failed analysis on clutch completion and species.}
#Transforming body mass data using log10 and square root to attempt to make data more normally distributed - unsuccessful.
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_log10()
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_sqrt()

ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_log10()
ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_sqrt()

ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_log10()
ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram() + scale_x_sqrt()

#Plotting a violin chart to explore whether body mass can explain successful clutch completion.
ggplot(data=penguins_data_clean, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="point", color="black")

#Plotting violin charts for each species to explore relationship between body mass and clutch completion in each species individually.
ggplot(data=adelie_data, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="pointrange", color="black")

ggplot(data=chinstrap_data, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="pointrange", color="black")

ggplot(data=gentoo_data, aes (x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color=species)) + scale_color_manual(values = species_colours) + labs (x= "Body mass (g)", y = "Clutch completion") + theme_bw() + stat_summary(species="mean", geom="pointrange", color="black")
```

```{r excess code which will probably be deleted -> unsure what is and is not useful and need to sort out}
#Plotting histogram to ensure data follows normal distribution
ggplot(data = penguins_data_clean, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = adelie_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = chinstrap_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))
ggplot(data = gentoo_data, aes(x=body_mass_g)) + geom_histogram(aes(color=species))

```{r possibly excess code relating to clutch completion vs body mass}
#Setting colour-blind friendly colour palette to distinguish data relating to different species, and defining shapes for each sex.
species_colours <- c("Adelie" = "darkorange", "Chinstrap" = "magenta2", "Gentoo" = "blue")
sex_shapes <- c("MALE" = "+", "FEMALE" = "20")

#Making function to create datasets only containing information on species, clutch completion, body mass and sex.
refining_penguins_data <- function(clean_data){
  clean_data %>% 
    select(species, body_mass_g, clutch_completion, sex) %>% 
    remove_NA()}

#Making new, refined dataset which only contains data on penguin species, size, sex and clutch completion.
penguins_data_refined <- refining_penguins_data(penguins_data_clean)
write.csv(penguins_data_refined, here("data", "penguins_data_refined.csv"))
penguins_data_refined <- read.csv(here("data","penguins_data_refined.csv"))
head(penguins_data_refined)

#Plotting violin chart of body mass against clutch completion for all species, without accounting for sex.
ggplot(data = penguins_data_refined, aes(x=body_mass_g, y=clutch_completion)) + geom_violin(aes(color = species, shape = sex), width = 1) + geom_jitter(aes(color = species), alpha = 0.25, width = 0.5) + scale_colour_manual(values = species_colours) + scale_shape_manual(values = sex_shapes)
```

```{r Code related to failed no interaction modelling for ANCOVA of culmen depth, length and species.}

plot(log_bill_model_no_int, which=2)
plot(log_bill_model_no_int, which=1)
summary(log_bill_model_no_int)
anova(log_bill_model_no_int)

plot_no_int <- ggplot(penguins_data_clean, aes(x=culmen_length_mm, y=culmen_depth_mm, color = species)) + geom_point() + geom_line() + scale_colour_manual(values = species_colours) + xlab("Culmen length (mm)") + ylab("Culmen depth (mm)") + theme_bw()

plot_no_int