---
title: "Reproducible Figures Assignment"
output:
  html_document: default
  pdf_document: default
date: "2023-11-29"
purpose: A script to analyse the Palmer Penguins dataset.
---

# Installing and loading packages

```{r}
#Loading packages needed for this project, which should all be in the RENV folder. Also loading functions to clean data from Palmer Penguins dataset, found in a separate folder.
library(tidyverse)
library(palmerpenguins)
library(janitor)
library(here)
library(ggplot2)
library(patchwork)
library(ragg)
library(svglite)

source(here("functions","cleaning_penguin_data.r"))
```

# Data Preparation:

```{r echo=FALSE}
#Preserving and loading raw data in its own .csv file.

write.csv(penguins_raw, here("data","penguins_raw.csv"))
penguins_data_raw<-read.csv(here("data","penguins_raw.csv"))


#Investigating raw data, including seeing what column names are included.

head(penguins_data_raw)
colnames(penguins_data_raw)


#Cleaning raw data to generate a cleaner dataset, using functions sourced from Lydia France's cleaning_penguins.r file.
penguins_data_clean <- cleaning_penguin_columns(penguins_data_raw)
colnames(penguins_data_clean)

#Writing and saving new .csv file for cleaned data, which can be found in the "data" folder. 
write.csv(penguins_data_clean, here("data","penguins_data_clean.csv"))
penguins_data_clean <- read.csv(here("data", "penguins_data_clean.csv"))
```

# QUESTION 1: Data Visualisation for Science Communication

## a) Creating a bad figure
Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. Do not make a boxplot.

```{r bad figure code, echo=FALSE}
#Creating a bad bar graph showing how body mass differs between islands to answer question 1 of assessment.

bad_bargraph_island_png <- ggplot(penguins_data_clean) +
                           geom_col(aes(x=island, y=body_mass_g))
print(bad_bargraph_island_png)

#Saving the misleading plot as a .png file in the "figures" folder.

agg_png("figures/bad_bargraph_island_bodymass.png", 
        width = 20,
        height = 20,
        units = "cm",
        res = 200,
        scaling = 0.5)

print(bad_bargraph_island_png)
dev.off()
```

## b) Why is this misleading?
Write about how your design choices mislead the reader about the underlying data (200-300 words). Include references.

My graph is correct, but some design choices mislead the reader.

Firstly, there is no information about each data point, such as the species of penguin each sample is a member of. This is misleading; penguin species is likely a confounding variable which impacts body mass, and the island(s) inhabited. This therefore may mislead a reader as the effects of different species are not explicitly shown, so a misleading correlation may be represented on my graph.

The lack of error bars is also misleading as it is unclear to a reader how much measurements for a certain treatment (in this case each island) vary. Therefore, it is unclear whether the samples taken from each island are representative of the population on each island or whether there is wide variation and improper sampling, or whether the data are precise and/or accurate.

On this graph, I also only show a summary of the data, in this case maximum body mass in each category (i.e on each island). This is misleading because important information may be hidden. For example, the maximum body mass for each island might be an outlier. This is also represented in the axis labels, which do not state maximum body mass but instead state only "body_mass_g".

Moreover, axis labels as specified in the .png plot are also very small, making it hard for a reader to understand what the graph is showing. Readers must look very closely at the axis labels to see the variables being compared. This is misleading because by not clearly showing your axis labels and any units associated with them, readers are more likely to miss key details about exactly what is being compared or what relationship is suggested by a graph between, which may cause them to believe a false correlation is present.

# QUESTION 2: Data Pipeline

## Introduction

Many factors can impact beak morphology in birds, including foraging and thermoregulation. It has been shown that in many bird families, different species which inhabit different niches and have different foraging behaviours tend to evolve different bill shapes, which impacts variables like culmen length and depth (Friedman, 2019; Çakar, 2024). This pattern has been observed and tested in the birds of prey, but not in penguins (Çakar, 2024). Therefore it will be interesting to investigate this in penguins and assess whether beak morphology does differ significantly between species.

Whether sex affects beak morphology is also an interesting question. It is known that species in the genus Pygoscelis display subtle sexual dimorphism in features like body mass, flipper length and bill size, with females usually being smaller than males. There is a possibility this will be reflected in the bill morphology of these penguins as well, with sex exerting impacts on bill shape as well as bill size (Gorman, 2014; Polito, et al., 2012). To study this in detail, this analysis will focus on only one penguin species (the gentoo penguin) to ensure only sex is examined, not the effect of species (Gorman, 2014).

This analysis, using data from the Palmer Penguins dataset, will test the relationship between culmen depth and culmen length in the penguin genus Pygoscelis. This analysis will assess the effect of species on this relationship, as well as the effect of sexual dimorphism in one penguin species, the gentoos (Pygoscelis papua). 

```{r Data Preparation- taken from lesson materials provided by Lydia France, University of Oxford, 2024}
#All packages used can be found at the top of the document.

#These lines of code preserve raw data in a separate csv file so it is not lost and can be easily recovered.
write.csv(penguins_raw, here("data","penguins_raw.csv"))
penguins_data_raw<-read.csv(here("data","penguins_raw.csv"))


#Investigating data- allows initial analysis of what data were collected and what values have been recorded for each.

head(penguins_data_raw)
colnames(penguins_data_raw)

#Cleaning raw data to generate a cleaner dataset- removing superfluous variables and making data more easily read by R and by humans. Code for cleaning_penguin_columns function can be found in "functions/cleaning_penguin_data.r".
penguins_data_clean<-cleaning_penguin_columns(penguins_data_raw)
colnames(penguins_data_clean)

#Writing new .csv file for new, clean dataset to keep this safe in case any accidental changes are made or code breaks, to allow code to be easily recovered. 
write.csv(penguins_data_clean, here("data","penguins_data_clean.csv"))
penguins_data_clean <- read.csv(here("data", "penguins_data_clean.csv"))
```

```{r Investigating all species data.}

#Creating histograms of data on culmen length in all species to investigate data and assess distribution of data points, as well as effects of any transformations.
ggplot(data=penguins_data_clean, aes(x=culmen_length_mm)) + geom_histogram()

#Log transforming culmen_length_mm using log10() function, and plotting histogram.
log_length <- log10(penguins_data_clean$culmen_length_mm) 
ggplot(data=penguins_data_clean, aes(x=log_length)) + geom_histogram()

#Square-root transforming culmen_length_mm using sqrt() function, and plotting histogram.
sqrt_length <- sqrt(penguins_data_clean$culmen_length_mm)
ggplot(data = penguins_data_clean, aes(x=sqrt_length)) + geom_histogram()
#Transformations were not that effective, although log transformation did make data slightly less right-skewed; will use log-transformed culmen length data for analysis.

#Creating histograms of data on culmen depth in all species to investigate data and assess distribution of data points, as well as effects of any transformations.
ggplot(data=penguins_data_clean, aes(x=culmen_depth_mm)) + geom_histogram()

#Log transforming culmen_depth_mm using log10() function, and plotting histogram.
log_depth <- log10(penguins_data_clean$culmen_depth_mm)
ggplot(data=penguins_data_clean, aes(x=log_depth)) + geom_histogram()

#Square-root transforming culmen_depth_mm using sqrt() function, and plotting histogram.
sqrt_depth <- sqrt(penguins_data_clean$culmen_depth_mm)
ggplot(data = penguins_data_clean, aes(x=sqrt_depth)) + geom_histogram()
#Transformations were not effective. Will use untransformed culmen depth data for further analysis.
```

```{r Creating exploratory figure}
#Creating an exploratory figure looking at the relationship between culmen depth and culmen length, as well as how this differs between species. 
#  scale_colour_manual function allows for different species to be designated the same colours    every time in a colour blind friendly palette, making the figure more accessible and more       reproducible. 
#     Also allows for more in depth exploration of initial data before statistical analysis          conducted.

#Making a colour blind-friendly palette useful for distinguishing species on figures. Specifying this ensures the different species are represented by the consistent colours on any graph, making this document more accessible (colour blind friendly) and more reproducible.
species_colours <- c("Adelie" = "darkorange", 
                     "Chinstrap" = "magenta2", 
                     "Gentoo" = "blue")

#Encodes information for all-species exploratory graph.
# scale_colour_manual() ensures each species is represented by the colours specified above in species_colours.
bill_size_exploratory <- ggplot(data=penguins_data_clean, 
                         aes(x=species, 
                            y=culmen_depth_mm)) + 
                         geom_point(aes(color=species)) + 
                         geom_jitter(aes(color = species), alpha = 0.25, position =                                     position_jitter(width = 0.4, seed=0)) +
                         scale_colour_manual(values = species_colours) + 
                         xlab("Culmen Length (mm)") + 
                         ylab("Culmen Depth (mm)")

#plot() function allows graph to be outputted.
plot(bill_size_exploratory)

#agg_png() saves figure as .png file in figures folder.
agg_png("figures/bill_size_exploratory.png", 
        width = 30,
        height = 15,
        units = "cm",
        res = 300,
        scaling = 1.125)

#print() ensures figure saved in "figures" folder actually has graph encoded within it, as opposed to blank file.
print(bill_size_exploratory)
dev.off()
```

```{r Sexual dimorphism}
#Decided to look at effects of sexual dimorphism, so encoded separate graph which codes for identification of sex as well as species through different shapes.
# scale_shape_manual allows specification of shape so different sexes can be designated the      same shapes consistently making the figure more reproducible. 
#   Also allows for more in depth exploration of initial data before statistical analysis          conducted.
#Plotted a scatter plot as comparing two numerical variables (culmen length and depth).
bill_size_sex_exploratory <- ggplot(data=penguins_data_clean, 
                         aes(x=species, 
                            y=culmen_depth_mm)) + 
                         geom_point(aes(color=species, shape=sex)) + 
                         geom_jitter(aes(color = species), alpha = 0.25, position =                                     position_jitter(width = 0.4, seed=0)) +
                         scale_colour_manual(values = species_colours) + 
                         scale_shape_manual(values = c(16, 17)) +
                         xlab("Culmen Length (mm)") + 
                         ylab("Culmen Depth (mm)")

#plot() function allows graph to be outputted.
plot(bill_size_sex_exploratory)

#Saving figure as .png file in figures folder.
agg_png("figures/bill_size__sex_exploratory.png", 
        width = 30,
        height = 15,
        units = "cm",
        res = 300,
        scaling = 1.125)

#Ensures figure saved in "figures" folder actually has graph encoded within it, as opposed to blank file.
print(bill_size_sex_exploratory)
dev.off()
```

```{r Gentoo specific}
#Subsetting data to only include that on gentoo penguins.
gentoo_data <- penguins_data_clean %>% 
               filter(species == "Gentoo") %>% 
               remove_NA()

#Creating histograms of data on culmen length in all Gentoos to investigate data and assess distribution of data points, as well as effects of any transformations.
#Histogram for untransformed data:
ggplot(data=gentoo_data, aes(x=culmen_length_mm)) + geom_histogram()

#Log transforming culmen_length_mm using log10() function, and plotting histogram.
gentoo_log_length <- log10(gentoo_data$culmen_length_mm)
ggplot(data=gentoo_data, aes(x=gentoo_log_length)) + geom_histogram()

#Square-root transforming culmen_depth_mm using sqrt() function, and plotting histogram.
gentoo_sqrt_length <- sqrt(gentoo_data$culmen_length_mm)
ggplot(data = gentoo_data, aes(x=gentoo_sqrt_length)) + geom_histogram()
```

```{r Creating exploratory figure for gentoo penguins in particular}
#Made exploratory figure to look more closely at one species in particular.
#Plotted a scatter plot as comparing two numerical variables (culmen length and depth).

#Making a colour blind-friendly palette useful for distinguishing sexes on figures. Specifying this ensures each sex is consistently represented by the same colours on any graph where sex is shown, making this document more accessible (as it is colour blind friendly) and more reproducible.
sex_colours <- c("FEMALE" = "purple",
                 "MALE" = "darkorange")

bill_size_exploratory_gentoo <- ggplot(data=gentoo_data, 
                         aes(x=species, 
                            y=culmen_depth_mm)) + 
                         geom_point(aes(color=sex, shape=sex)) + 
                         geom_jitter(aes(color = sex), alpha = 0.25, position =                                     position_jitter(width = 0.4, seed=0)) +
                         scale_colour_manual(values = sex_colours) + 
                         scale_shape_manual(values = c(16, 17)) +
                         xlab("Culmen Length (mm)") + 
                         ylab("Culmen Depth (mm)")

plot(bill_size_exploratory_gentoo)

#Saving gentoo figure as .png files in figures folder.
agg_png("figures/bill_size_gentoo_exploratory.png", 
        width = 30,
        height = 15,
        units = "cm",
        res = 300,
        scaling = 1.125)

print(bill_size_exploratory_gentoo)
dev.off()

```

## Hypotheses

### All species
Null hypotheses for species effect: 1: There is no effect of culmen length on culmen depth. 2: There is no effect of species on culmen depth. 3: There is no interaction between species and culmen length.

Alternative hypotheses for species effect: 1: There is a significant effect of culmen length on culmen depth. . 2: There is a significant effect of species on culmen depth. 3: There is a significant interaction between species and culmen length.

###Gentoo penguins
Null hypotheses for sex effect in Gentoos: 1: There is no effect of culmen length on culmen depth. 2: There is no effect of sex on culmen depth. 3: There is no interaction between sex and culmen length.

Alternative hypotheses for sex effect in Gentoos: 1: There is a significant effect of culmen length on culmen depth. . 2: There is a significant effect of species on culmen depth. 3: There is a significant interaction between species and culmen length.

## Statistical Methods

In this analysis, an ANCOVA test will be conducted, allowing for testing of the relationship between culmen depth and culmen length as well as culmen depth and species, alongside testing if there is a significant interaction between culmen length and species. An ANCOVA is suitable for this analysis as there is a numerical response variable (culmen depth), a numerical explanatory variable (culmen length) and a categorical explanatory variable (species).

In Gentoo penguins only, an ANCOVA test will also be conducted, testing the relationship between culmen depth and culmen length as well as culmen depth and sex, alongside testing if there is a significant interaction between culmen length and sex.

```{r Statistics- all species}

#Creating model including interaction term.
log_bill_model_int <- lm(culmen_depth_mm ~ log_length * species, penguins_data_clean)

#Makes plots to ensure assumptions of normality (specified by which=2) and homoscedasticity (specified by which=1) are not violated.
# par() function allows plotting of multi-panel figure.
par(mfrow = c(1,2))
plot(log_bill_model_int, which = 2)
plot(log_bill_model_int, which = 1)

#Checking outputs of analysis of model to assess adjusted R-squared and values for slope and intercept of each group.
summary(log_bill_model_int)

#Conducting an ANOVA of the above all-species model with interaction to assess what the significant effects are.
allspecies_int_anova <- aov(log_bill_model_int)
summary(allspecies_int_anova)

#Model without interaction term - necessary as no significant interaction found.
log_bill_model_noint <- lm(culmen_depth_mm ~ log_length + species, penguins_data_clean)

#Making plot to ensure assumptions of analysis are not violated.
par(mfrow = c(1,2))
plot(log_bill_model_noint, which = 2)
plot(log_bill_model_noint, which = 1)

#Checking outputs of analysis of model to assess adjusted R-squared and values for slope and intercept of each group.
summary(log_bill_model_noint)

#Conducting an ANOVA of the above all-species model without interaction to assess what the significant effects are.
allspecies_noint_anova <- aov(log_bill_model_noint)
summary(allspecies_noint_anova)

#Conducting ANOVA to assess if there is a significant interaction between species and culmen length (mm)
allspecies_test_fits_diff <- anova(log_bill_model_int, log_bill_model_noint)
allspecies_test_fits_diff
```

```{r Statistics- Gentoos}
#Creating gentoo-specific model including interaction term.
gentoo_log_bill_model_int <- lm(culmen_depth_mm ~ gentoo_log_length * sex, gentoo_data)

#Makes plots to ensure assumptions of analysis not violated.
par(mfrow = c(1,2))
plot(gentoo_log_bill_model_int, which = 2)
plot(gentoo_log_bill_model_int, which = 1)

#Checking outputs of analysis of model to assess adjusted R-squared and values for slope and intercept of each group.
summary(gentoo_log_bill_model_int)

#Conducting an ANOVA of the above gentoo-specific model to assess what the significant effects are.
gentoo_int_anova <- aov(gentoo_log_bill_model_int)
summary(gentoo_int_anova)

#Creating gentoo-specific model without interaction term.
gentoo_log_bill_model_noint <- lm(culmen_depth_mm ~ gentoo_log_length + sex, gentoo_data)

#Makes plots to ensure assumptions of analysis not violated.
par(mfrow = c(1,2))
plot(gentoo_log_bill_model_noint, which = 2)
plot(gentoo_log_bill_model_noint, which = 1)

#Checking outputs of analysis of model without interaction to assess adjusted R-squared and values for slope and intercept of each group.
summary(gentoo_log_bill_model_noint)

#Conducting an ANOVA of the gentoo-specific model without interaction to assess what the significant effects are.
gentoo_noint_anova <- aov(gentoo_log_bill_model_noint)
summary(gentoo_noint_anova)

#Conducting ANOVA to assess if there is a significant interaction between sex and culmen length in Gentoo penguins.
gentoo_test_fit_diff <- anova(gentoo_log_bill_model_int, gentoo_log_bill_model_noint)
gentoo_test_fit_diff
```

## Results & Discussion

Having conducted an ANCOVA analysis to test the effects of species and log culmen length (mm) on culmen depth (mm), it has been confirmed that while species and culmen length (mm) do both affect culmen depth (mm), there is no interaction between species and culmen length (mm). Therefore, while we fail to reject the null hypothesis that there is no interaction between species and culmen length, we can reject two null hypotheses: 1- culmen length does not affect culmen depth. 2- species does not affect culmen depth. The adjusted R-squared value of the model lm(culmen_depth \~ species \* culmen_length) was 0.7664, suggesting 76.64% of variance in the data can be explained by the model. The p-values for both effect of species and effect of log culmen length on culmen depth were both very small, and the F-values were both greater than 1 (496.1031 and 126.6277 respectively).

This aligns with what might be expected as increasing culmen length is likely to be correlated with increasing culmen depth, increasing the size of the beak overall. Moreover, as different species are different sizes and inhabit different ecological niches, they are likely to evolve different beak morphologies in response which are better suited to foods they hunt. That said, the fact there is no interaction between species and culmen length suggests...

An ANCOVA analysis was also conducted on the effects of sex and log culmen length on culmen depth. This analysis confirmed that while sex and culmen length (mm) do have a statistically significant effect on culmen depth (mm), there is again no interaction between the two explanatory variables. Therefore, we fail to reject the null hypothesis that there is no interaction between sex and culmen length, suggesting that sexual dimorphism is not at play here. This is surprising, given that on the exploratory figure showing data on gentoos, sex seemed to play a role in determining culmen depth with respect to culmen length, but this is not the case.

That said, two null hypotheses can be rejected: 1- culmen length does not affect culmen depth in gentoo penguins. 2- sex does not affect culmen depth in gentoo penguins. The adjusted R-squared value for the model lm(gentoo_culmen_depth \~ sex \* gentoo_log_culmen_length) was 0.6094, showing that 60.94% of variance recorded can be explained by the model put forward. The F-value of sex for this model was 171.592, and for log_length was 15.3785, but for the interaction an F-value of 0.2038 was recorded, which is less than 1 and therefore shows there is no significant interaction. While log_length has a statistically significant effect on culmen_depth in gentoo penguins, it has a lower F-value and p-value than sex, suggesting the relationship in this penguin species is weak.

```{r Plotting Results- All species}

#Making results plot, showing relationship between culmen depth, culmen length and species.
# geom_smooth() allows plotting of regression lines on the results figure.
all_species_plot_int <- ggplot(penguins_data_clean, 
                               aes(x=log_length, y=culmen_depth_mm, color = species)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_colour_manual(values = species_colours) + 
  xlab("Log culmen length (mm)") + 
  ylab("Culmen depth (mm)") + 
  theme_bw()

plot(all_species_plot_int)

agg_png("figures/all_species_results_plot.png", 
        width = 30,
        height = 15,
        units = "cm",
        res = 300,
        scaling = 1.125)

print(all_species_plot_int)
dev.off()

```

```{r}
#Making results plot, showing relationship between culmen depth, culmen length and sex in gentoos.
gentoo_plot_int <- 
ggplot(gentoo_data, 
       aes(x=gentoo_log_length, y=culmen_depth_mm, color = sex)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_colour_manual(values = sex_colours) + 
  xlab("Log Culmen length (mm)") + 
  ylab("Culmen depth (mm)") + 
  theme_bw()

plot(gentoo_plot_int)

agg_png("figures/gentoo_results_plot.png", 
        width = 30,
        height = 15,
        units = "cm",
        res = 300,
        scaling = 1.125)

print(gentoo_plot_int)
dev.off()
```

## Conclusion

##References

*You will be marked on the following:*

a)  Your code for readability and functionality
b)  Your figures for communication
c)  Your text communication of your analysis

# QUESTION 3: Open Science

## a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

## b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

## c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

## d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*
