---
title: "Reproducible Figures Assignment"
output:
  html_document: default
  pdf_document: default
date: "2023-11-29"
---

# Installing and loading packages

```{r echo=FALSE}
#Loading packages needed for this project, which should all be in the RENV folder. Also loading functions to clean data from Palmer Penguins dataset, found in a separate folder.
library(tidyverse)
library(palmerpenguins)
library(janitor)
library(here)
library(ggplot2)
library(patchwork)
library(ragg)
library(svglite)

source(here("functions","cleaning_penguin_data.r"))
```

# Data Preparation:

```{r echo=FALSE}
#Preserving and loading raw data in its own .csv file.

write.csv(penguins_raw, here("data","penguins_raw.csv"))
penguins_data_raw<-read.csv(here("data","penguins_raw.csv"))


#Investigating raw data, including seeing what column names are included.

head(penguins_data_raw)
colnames(penguins_data_raw)


#Cleaning raw data to generate a cleaner dataset, using functions sourced from Lydia France's cleaning_penguins.r file.
penguins_data_clean <- cleaning_penguin_columns(penguins_data_raw)
colnames(penguins_data_clean)

#Writing and saving new .csv file for cleaned data, which can be found in the "data" folder. 
write.csv(penguins_data_clean, here("data","penguins_data_clean.csv"))
penguins_data_clean <- read.csv(here("data", "penguins_data_clean.csv"))
```

## QUESTION 1: Data Visualisation for Science Communication

### a) Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. Do not make a boxplot.

```{r bad figure code, echo=FALSE}
#Creating a bad bar graph showing how body mass differs between islands, to answer question 1 of assessment.

bad_bargraph_island_png <- ggplot(penguins_data_clean) +
                           geom_col(aes(x=island, y=body_mass_g))
print(bad_bargraph_island_png)

#Saving the misleading plot as a .png file in the "figures" folder.

agg_png("figures/bad_bargraph_island_bodymass.png", 
        width = 20,
        height = 20,
        units = "cm",
        res = 200,
        scaling = 0.5)

print(bad_bargraph_island_png)
dev.off()
```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words). Include references.

My graph is correct but some design choices mislead the reader.

Firstly, there is no information about each data point, such as the species of penguin each sample is a member of. This is misleading; penguin species is likely a confounding variable which impacts penguin size and the island(s) inhabited. This therefore may mislead a reader as the effects of different species are not explicitly shown, so a misleading correlation may be represented on my graph. 

The lack of error bars is also misleading as it is unclear to a reader how much measurements for a certain treatment (in this case each island) vary. Therefore, it is unclear whether the samples taken from each island are representative of the population on each island or whether there is wide variation and improper sampling, or whether the data are precise and/or accurate.

On this graph, I also only show a summary of the data, in this case maximum body mass in each category (i.e on each island). This is misleading because important information may be hidden. For example, the maximum body mass for each island might be an outlier. This is also represented in the axis labels, which do not state maximum body mass but instead state only "body_mass_g". 

Moreover, axis labels as specified in the .png plot are also very small, making it hard for a reader to understand what the graph is showing. Readers must look very closely at the axis labels to see the variables being compared. This is misleading because by not clearly showing your axis labels and any units associated with them, readers are more likely to miss key details about exactly what is being compared or what relationship is suggested by a graph between, which may cause them to believe a false correlation is present.


## Question 2

### Introduction

In this analysis, I will assess whether there is a relationship between culmen (beak) depth and culmen length, and whether species significantly affects this.

```{r Data Preparation- taken from lesson materials provided by Lydia France, University of Oxford, 2024}
#All packages used can be found at the top of the document.

#These lines of code preserve raw data in a separate csv file so it is not lost and can be easily recovered.
write.csv(penguins_raw, here("data","penguins_raw.csv"))
penguins_data_raw<-read.csv(here("data","penguins_raw.csv"))


#Investigating data- allows initial analysis of what data were collected and what values have been recorded for each.

head(penguins_data_raw)
colnames(penguins_data_raw)


#Cleaning raw data to generate a cleaner dataset- removing superfluous variables and making data more easily read by R and by humans. Code for cleaning_penguin_columns function can be found in "functions/cleaning_penguin_data.r".
penguins_data_clean<-cleaning_penguin_columns(penguins_data_raw)
colnames(penguins_data_clean)

#Writing new .csv file for new, clean dataset to keep this safe in case any accidental changes are made or code breaks, to allow code to be easily recovered. 
write.csv(penguins_data_clean, here("data","penguins_data_clean.csv"))
penguins_data_clean <- read.csv(here("data", "penguins_data_clean.csv"))
```

```{r Creating an exploratory figure.}
#Making a colour blind-friendly palette useful for distinguishing species on figures.
species_colours <- c("Adelie" = "darkorange", 
                     "Chinstrap" = "magenta2", 
                     "Gentoo" = "blue")

#Creating histograms of culmen length to investigate data and assess distribution of data points, as well as effects of any transformations.
ggplot(data=penguins_data_clean, aes(x=culmen_length_mm)) + geom_histogram()
log_length <- log10(penguins_data_clean$culmen_length_mm)
ggplot(data=penguins_data_clean, aes(x=log_length)) + geom_histogram()
sqrt_length <- sqrt(penguins_data_clean$culmen_length_mm)
ggplot(data = penguins_data_clean, aes(x=sqrt_length)) + geom_histogram()
#Transformations were not that effective, although log transformation did make data slightly less right-skewed. Will consider using log_length variable for analysis.

#Creating histograms of culmen depth to investigate data and assess distribution of data points, as well as effects of any transformations.
ggplot(data=penguins_data_clean, aes(x=culmen_depth_mm)) + geom_histogram()
log_depth <- log10(penguins_data_clean$culmen_depth_mm)
ggplot(data=penguins_data_clean, aes(x=log_depth)) + geom_histogram()
sqrt_depth <- sqrt(penguins_data_clean$culmen_depth_mm)
ggplot(data = penguins_data_clean, aes(x=sqrt_depth)) + geom_histogram()
#Transformations were not effective. Will use culmen depth data as is for further analysis.

#Creating an exploratory figure looking at the relationship between beak depth and beak length, as well as how this differs between species. 
#  scale_colour_manual function allows for different species to be designated the same colours    every time in a colour blind friendly palette, making the figure more acessible and more       reproducible. Also allows for more in depth exploration of initial data before statistical     analysis conducted.
#  Plotted a scatter plot for this figure as comparing two numerical variables.
bill_size_exploratory <- ggplot(data=penguins_data_clean, 
                         aes(x=culmen_length_mm, 
                            y=culmen_depth_mm)) + 
                         geom_point(aes(color=species)) + 
                         scale_colour_manual(values = species_colours) + 
                         xlab("Culmen Length (mm)") + 
                         ylab("Culmen Depth (mm)")

plot(bill_size_exploratory)

#Saving figure as .png file in functions folder.
agg_png("figures/bill_size_exploratory.png", 
        width = 30,
        height = 15,
        units = "cm",
        res = 300,
        scaling = 1.125)

print(bill_size_exploratory)
dev.off()

```


### Hypothesis

Null hypotheses:
1: There is no effect of culmen length on culmen depth. .
2: There is no effect of species on culmen depth.
3: There is no interaction between species, culmen length or culmen depth.

Alternative hypotheses:
1: There is a significant effect of culmen length on culmen depth. .
2: There is a significant effect of species on culmen depth.
3: There is a significant interaction between species, culmen length and culmen depth.


### Statistical Methods

To test whether culmen length can predict culmen depth, and whether this is affected by species, I will use an ANCOVA test.

```{r Statistics}

log_bill_model_int <- lm(culmen_depth_mm ~ species * log_length, penguins_data_clean)

plot(log_bill_model_int, which = 2)
plot(log_bill_model_int, which = 1)

summary(log_bill_model_int)

anova(log_bill_model_int)

plot_int <- ggplot(penguins_data_clean, aes(x=culmen_length_mm, y=culmen_depth_mm, color = species)) + geom_point() + geom_smooth(method = "lm") + scale_colour_manual(values = species_colours) + xlab("Culmen length (mm)") + ylab("Culmen depth (mm)") + theme_bw()

plot_int

log_bill_model_no_int <- lm(culmen_depth_mm ~ species + log_length, data =  penguins_data_clean)

plot(log_bill_model_no_int, which=2)
plot(log_bill_model_no_int, which=1)
summary(log_bill_model_no_int)
anova(log_bill_model_no_int)

plot_no_int <- ggplot(penguins_data_clean, aes(x=culmen_length_mm, y=culmen_depth_mm, color = species)) + geom_point() + geom_line(aes()) + scale_colour_manual(values = species_colours) + xlab("Culmen length (mm)") + ylab("Culmen depth (mm)") + theme_bw()

plot_no_int

```

### Results & Discussion

```{r Plotting Results}

# Make sure your code prints. 

```

### Conclusion

*You will be marked on the following:*

a)  Your code for readability and functionality
b)  Your figures for communication
c)  Your text communication of your analysis

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*
