---
title: "Reproducible Figures Assignment"
output:
  html_document: default
  pdf_document: default
date: "2023-11-29"
---

# Installing and loading packages

```{r echo=FALSE}
library(tidyverse)
library(palmerpenguins)
library(janitor)
library(here)
library(ggplot2)
library(patchwork)
library(ragg)
library(svglite)

source(here("functions","cleaning_penguin_data.r"))
```

# Data Preparation:

```{r echo=FALSE}
#Preserving and loading raw data.

write.csv(penguins_raw, here("data","penguins_raw.csv"))
penguins_data_raw<-read.csv(here("data","penguins_raw.csv"))


#Investigating data

head(penguins_data_raw)
colnames(penguins_data_raw)


#Cleaning raw data to generate a cleaner dataset.

cleaning_penguin_columns <- function(raw_data){
    raw_data %>% 
    clean_names() %>% 
    shorten_species() %>% 
      remove_empty(c("rows", "cols")) %>% 
      select(-starts_with("delta")) %>% 
      select(-comments)}
      
colnames(penguins_data_raw)
penguins_data_clean<-cleaning_penguin_columns(penguins_data_raw)
colnames(penguins_data_clean)

#writes new csv for clean penguins data. 
write.csv(penguins_data_clean, here("data","penguins_data_clean.csv"))
```

## QUESTION 1: Data Visualisation for Science Communication

### a) Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. Do not make a boxplot.

```{r bad figure code, echo=FALSE}
bad_bargraph_island_png<-ggplot(penguins_data_clean) + geom_col(aes(x=island, y=body_mass_g))

ggplot(penguins_data_clean) + geom_col(aes(x=island, y=body_mass_g))

agg_png("figures/bad_bargraph_island_bodymass.png", 
        width = 20,
        height = 20,
        units = "cm",
        res = 200,
        scaling = 0.5)

print(bad_bargraph_island_png)
dev.off()

bad_bargraph_clutch_png<-ggplot(penguins_data_clean) + geom_col(aes(x=clutch_completion, y=body_mass_g))

agg_png("figures/bad_bargraph_clutch_bodymass.png", 
        width = 20,
        height = 20,
        units = "cm",
        res = 200,
        scaling = 0.5)

print(bad_bargraph_clutch_png)
dev.off()

bad_scatterplot_png<-ggplot(penguins_data_clean, aes(x=body_mass_g, y=culmen_length_mm)) + geom_point()

agg_png("figures/bad_scatterplot_bodymass_culmenlength.png", 
        width = 20,
        height = 20,
        units = "cm",
        res = 200,
        scaling = 0.5)

print(bad_scatterplot_png)
dev.off()

ggplot(penguins_data_clean) + geom_col(aes(x=island, y=body_mass_g, color=species))

island_bm_anova <- aov(body_mass_g ~ island + species, data=penguins_data_clean)
summary(island_bm_anova)
```

### b) Write about how your design choices mislead the reader about the underlying data (200-300 words). Include references.

My graph is correct but misleads the reader in a few ways. 

To start with, there is no information about any confounding variables which could better explain a supposed relationship between island and body mass. For example, the 3 penguin species included in the Palmer Penguins dataset (Adelie, Gentoo and Chinstrap) appear to inhabit different islands. This means that although there could be a statistically significant relationship between island and body mass, the effect of different species found on each island has not been accounted for. 

No measures of statistical significance (such as asterisks signifying significant results) have been shown on this graph either, meaning that confounding variables aside, it is unclear to the reader whether there is a significant difference in penguin body mass between islands studied or not.

Moreover, my graph is hard to read. On this graph, the axes labels are very small, so it is difficult for a reader to pick out the explanatory and response variables and therefore the results the graph is trying to communicate. The labels themselves are also subpar, and do not adequately explain the contents of the graph; the y-axis label only states it is measuring “body_mass_g”, but the information contained in the graphs current form is the maximum body mass recorded on each island. Each bar on the bar graph is the same colour, which makes it harder to differentiate the categories and so makes this graph even more unclear. 


## Question 2

### Introduction

```{r Data Exploration}
library(palmerpenguins)

write.csv(penguins_raw, here("data","penguins_raw.csv"))
penguins_data_raw<-read.csv(here("data","penguins_raw.csv"))


#Investigating data

head(penguins_data_raw)
colnames(penguins_data_raw)


#Cleaning raw data to generate a cleaner dataset.

cleaning_penguin_columns <- function(raw_data){
    raw_data %>% 
    clean_names() %>% 
    shorten_species() %>% 
      remove_empty(c("rows", "cols")) %>% 
      select(-starts_with("delta")) %>% 
      select(-comments)}
      
colnames(penguins_data_raw)
penguins_data_clean<-cleaning_penguin_columns(penguins_data_raw)
colnames(penguins_data_clean)

#writes new csv for clean penguins data. 
write.csv(penguins_data_clean, here("data","penguins_data_clean.csv"))
```

```

### Hypothesis

### Statistical Methods

```{r Statistics}

# Make sure your code prints. 

```

### Results & Discussion

```{r Plotting Results}

# Make sure your code prints. 

```

### Conclusion

*You will be marked on the following:*

a)  Your code for readability and functionality
b)  Your figures for communication
c)  Your text communication of your analysis

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:*

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*
